---
title: "GBIF Biotic Interactions Data - Taxonomic Analysis"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages

```{r library}
library(dplyr)
library(viridis)
library(scales)
library(hrbrthemes)
```

## Load and prepare Data

Load all interaction records
```{r interaction-data}
interactions.all <- read.csv('data/interactions-taxon-loc-time.tsv.gz', sep = '\t', quote="")
head(interactions.all)
dim(interactions.all)
```

Replace kingdom with superkingdom where kingdom is empty
```{r}
interactions <- interactions.all %>%
  mutate(
    sourceTaxonKingdomName=ifelse(sourceTaxonKingdomName == "" & sourceTaxonSuperkingdomName!="", sourceTaxonSuperkingdomName, sourceTaxonKingdomName),
    targetTaxonKingdomName=ifelse(targetTaxonKingdomName == "" & targetTaxonSuperkingdomName!="", targetTaxonSuperkingdomName, targetTaxonKingdomName)
    )
```

Remove interactions without taxonomic identification
```{r}
interactions.excluded <- interactions %>%
  filter(sourceTaxonId == 'no:match' | targetTaxonId == 'no:match') %>%
  filter(sourceTaxonKingdomName=="" | targetTaxonKingdomName=="")

interactions <- interactions %>%
  filter(sourceTaxonId != 'no:match' & targetTaxonId != 'no:match') %>%
  filter(sourceTaxonKingdomName!="" & targetTaxonKingdomName!="")
dim(interactions)
```

Examples of names excluded
```{r}
interactions.excluded %>%
  filter(sourceTaxonKingdomName=="") %>%
  select(sourceTaxonKingdomName, sourceTaxonName) %>%
  dplyr::rename(kingdomName=sourceTaxonKingdomName, taxonName=sourceTaxonName) %>%
  bind_rows(
    interactions.excluded %>%
      filter(targetTaxonKingdomName=="") %>%
      select(targetTaxonKingdomName, targetTaxonName) %>%
      dplyr::rename(kingdomName=targetTaxonKingdomName, taxonName=targetTaxonName) 
  ) %>%
  pivot_longer(cols=c(taxonName)) %>%
  group_by(value) %>%
  dplyr::summarize(n=n()) %>%
  ungroup() %>%
  mutate(total=sum(n), perc=round(n/total*100,2)) %>%
  arrange(desc(n))
```

Counting interactions by type
```{r}
interactions.all %>%
  group_by(interactionTypeName) %>%
  dplyr::summarise(n=n(),perc=round(n/nrow(.)*100,2)) %>%
  arrange(desc(n))
```
Create a list of taxon names from interaction records:

```{r}
my.species <- interactions %>%
  dplyr::rename(id=sourceTaxonId,speciesName=sourceTaxonSpeciesName,kingdomId=sourceTaxonKingdomId,kingdomName=sourceTaxonKingdomName) %>%
  select(id,speciesName,kingdomId,kingdomName) %>%
  bind_rows(
    interactions %>%
      dplyr::rename(id=targetTaxonId, speciesName=targetTaxonSpeciesName, kingdomId=targetTaxonKingdomId, kingdomName=targetTaxonKingdomName) %>%
      select(id,speciesName,kingdomId,kingdomName)
  ) %>%
  dplyr::distinct(id,speciesName,kingdomId,kingdomName) 
head(my.species)
dim(my.species)
```

NCBI uses IPTV for virus names but here we grouped all names into the same kingdom: Viruses. Define a list of virus names:

```{r viruses}
viruses <- c('Riboviria', 'Monodnaviria', 'Naldaviricetes', 'Duplodnaviria', 'Pospiviroidae',
             'Varidnaviria', 'Adnaviria', 'Polydnaviridae', 'Alphasatellitidae',
             'Tolecusatellitidae', 'Dinodnavirus', 'Thaspiviridae', 'Rhizidiovirus',
             "Portogloboviridae","Ovaliviridae", "Guttaviridae", "Fuselloviridae",    
             "Anelloviridae", "Avsunviroidae", "Bicaudaviridae", "Spiraviridae",
             "Clavaviridae", "Finnlakeviridae", "Halspiviridae", "Globuloviridae",
             "Ampullaviridae", "Ribozyviria", "Plasmaviridae", 'Orthornavirae',
             'Shotokuvirae', 'Heunggongvirae', 'Pararnavirae', 'Bamfordvirae',
             'Helvetiavirae', 'Sangervirae', 'Loebvirae', 'Zilligvirae', 'Trapavirae')
```

Counting interactions by kingdom
```{r}
interactions %>%
  dplyr::rename(id=sourceTaxonId,speciesName=sourceTaxonSpeciesName,kingdomId=sourceTaxonKingdomId,kingdomName=sourceTaxonKingdomName) %>%
  select(id,speciesName,kingdomId,kingdomName) %>%
  bind_rows(
    interactions %>%
      dplyr::rename(id=targetTaxonId, speciesName=targetTaxonSpeciesName, kingdomId=targetTaxonKingdomId, kingdomName=targetTaxonKingdomName) %>%
      select(id,speciesName,kingdomId,kingdomName)
  ) %>%
  mutate(kingdomName=ifelse(kingdomName %in% viruses, "Viruses", kingdomName)) %>%
  mutate(kingdomName=ifelse(kingdomName %in% c('Plantae','Viridiplantae','Chloroplastida','Archaeplastida'), 'Plant/Viridiplantae/Chloroplastida/Archaeplastida', kingdomName)) %>%
  mutate(kingdomName=ifelse(kingdomName %in% c('Animalia', 'Metazoa', 'Protozoa'), 'Animalia/Metazoa/Protozoa', kingdomName)) %>%
  mutate(kingdomName=ifelse(kingdomName %in% c('Fungi', 'Chromista'), 'Fungi/Chromista', kingdomName)) %>%
  mutate(kingdomName=ifelse(kingdomName == 'Eukaryota', 'incertae sedis', kingdomName)) %>%
  group_by(kingdomName) %>%
  dplyr::summarise(n=n(), perc=round(n/(nrow(interactions)*2)*100,1)) %>%
  arrange(desc(n))
 
```

Loading GBIF Backbone Taxonomy

```{r gbif-data}
gbif.species <- read.csv('data/gbif-species-list.tsv', header = F, sep='\t')
names(gbif.species) <- c('id', 'rank', 'kingdomId', 'scientificName', 'speciesName')
head(gbif.species)
dim(gbif.species)
```

Loading GBIF Kingdom IDs

```{r}
kingdoms <-read.csv('data/gbif-kingdom.tsv', header = F, sep = '\t', col.names=c('kingdomId','kingdomName'))
kingdoms
```

Select only GBIF taxon names and remove the `GBIF:` prefix from taxon Id's

```{r}
my.gbif.species <- my.species %>%
  filter(stringr::str_detect(id,"GBIF:")) %>%
  mutate(id = stringr::str_replace_all(id,'GBIF:', ''), kingdomId = stringr::str_replace(kingdomId, 'GBIF:', ''))
head(my.gbif.species)
```

Join data.frames into one and count number of species names by kingdom

```{r}
species.by.kingdom <- gbif.species %>%
  dplyr::count(kingdomId, name='gbif') %>%
  left_join(kingdoms, by = join_by(kingdomId)) %>%
  select(kingdomName,gbif) %>%
  dplyr::left_join(my.gbif.species %>% dplyr::count(kingdomName, name='my'), by = join_by(kingdomName)) %>%
  filter(kingdomName != 'Viruses' & kingdomName != 'incertae sedis')
species.by.kingdom$my[is.na(species.by.kingdom$my)] <- 0
species.by.kingdom
```

Reorganize data.frame to have counts by source (GBIF, my)

```{r}
data.gbif <- species.by.kingdom %>% gather(key = "source", value="value", -c(1)) %>%
  bind_rows(group_by(., source) %>% dplyr::summarise(value=sum(value)) %>% mutate(kingdomName='All Kingdoms'))
head(data.gbif)
```

Loading NCBI Taxonomy

```{r ncbi}
ncbi.taxonomy <- read.csv('data/ncbi-taxonomy-reformatted.tsv', header = F, sep='\t', quote = "")
names(ncbi.taxonomy) <- c('id','v2','superKingdom', 'kingdom', 'phylum', 'class', 'order', 'family', 'genus', 'species')
head(ncbi.taxonomy)
```

Clean data: remove double quotes in names, set super kingdom as kingdom where kingdom is missing (`NA` or `""`) and change kingdom to Viruses if taxon name is in `viruses` list

```{r}
ncbi.taxonomy <- ncbi.taxonomy %>%
  select(-v2) %>%
  mutate(kingdom=stringr::str_trim(kingdom), superKingdom=stringr::str_trim(superKingdom)) %>%
  mutate(kingdomNorm=ifelse(kingdom == "" & superKingdom != '', superKingdom, kingdom)) %>%
  mutate(kingdomNorm=ifelse(kingdom %in% viruses, "Viruses", kingdomNorm)) %>%
  mutate(kingdomNorm = stringr::str_remove_all(kingdomNorm, '"'))
head(ncbi.taxonomy)
```

Loading DwC-A list of taxon names

```{r}
my.ncbi.species <- my.species %>%
  filter(stringr::str_detect(id,"NCBI:")) %>%
  mutate(kingdomName=ifelse(kingdomName %in% viruses, "Viruses", kingdomName)) %>%
  mutate(
    id = stringr::str_replace_all(id,'NCBI:', ''),
    kingdomId = stringr::str_replace_all(kingdomId, 'NCBI:', '')
  )
head(my.ncbi.species)
dim(my.ncbi.species)
```

Clean data: remove double quotes in names and set super kingdom as kingdom where kingdom is missing (`NA` or `""`)

```{r}
my.ncbi.species <- my.ncbi.species %>%
  mutate(across(is.character, ~ stringr::str_remove_all('"')))
head(my.ncbi.species)
```

Join data.frames

```{r}
species.ncbi.by.kingdom <- ncbi.taxonomy %>%
  dplyr::count(kingdomNorm, name='ncbi') %>%
  select(kingdomNorm,ncbi) %>%
  mutate(kingdomNorm=stringr::str_trim(kingdomNorm)) %>%
  left_join(my.ncbi.species %>% dplyr::count(kingdomName, name='my') %>% mutate(kingdomName = stringr::str_trim(kingdomName)), by = join_by(kingdomNorm==kingdomName))
```

Set count to `0` if it is `NA`

```{r}
species.ncbi.by.kingdom$my[is.na(species.ncbi.by.kingdom$my)] <- 0
species.ncbi.by.kingdom
```

Reorganize data

```{r}
data.ncbi <- species.ncbi.by.kingdom %>% gather(key = "source", value="value", -c(1)) %>%
  bind_rows(group_by(., source) %>% dplyr::summarise(value=sum(value)) %>% mutate(kingdomNorm='All Kingdoms'))
```

Reorder factors:

```{r}
data.ncbi <- data.ncbi %>%
  filter(kingdomNorm!='') |>
  group_by(kingdomNorm, source) |> 
  mutate(source=as.factor(source))
```

## Taxonomic Coverage

### Plot Taxonomic coverage by kingdom

GBIF

```{r plot}
plt <- data.gbif |> 
  group_by(kingdomName, source) |> 
  ggplot(aes(x = source, y = value, fill=kingdomName)) +
  geom_col() +
  scale_fill_viridis(
    discrete = T,
    option='A',
    guide='none'
  ) +
  scale_y_continuous(labels = label_number(suffix = " M", scale = 1e-6), n.breaks = 4) + # millions
  scale_x_discrete(labels=c('GBIF', 'Local')) + 
  facet_wrap(~ kingdomName, scales = 'free') + 
  coord_flip() + 
  labs(fill='Kingdom', title = "Taxonomic coverage by kingdom", subtitle =expression("Comparison of the number of species names found in \n analyzed data (local) and in GBIF Backbone Taxonomy")) +
  theme_ipsum(
    base_family = 'Arial',
  ) +
  theme(
    panel.spacing = unit(0,'lines'),
    plot.subtitle = element_text(margin = margin(2,0,0,0, "mm")),
    axis.title = element_blank(),
    strip.text = element_text(size=10),
    axis.text.x = element_text(size=6),
    axis.text.y = element_text(size=8),
  ) 
plt
```

Save plot:

```{r saveplot}
ggsave('images/taxon-cov-by-kingdom-gbif.png', plt, dpi = 300)
```

NCBI

```{r}
plt <- data.ncbi %>%
  filter(kingdomNorm!='') |>
  group_by(kingdomNorm, source) |> 
  mutate(source=as.factor(source)) %>%
  ggplot(aes(x = source, y = value, fill=kingdomNorm)) +
  geom_col() +
  scale_fill_viridis(
    discrete = T,
    option='A',
    guide='none'
  ) +
  scale_y_continuous(labels = label_number(suffix = " M", scale = 1e-6), n.breaks = 4) + # millions
  scale_x_discrete(labels=c('NCBI', 'Local'), limits=rev(levels(data.ncbi$source))) + 
  facet_wrap(~ kingdomNorm, scales = 'free') + 
  coord_flip() + 
  labs(fill='Kingdom', title = "Taxonomic coverage by kingdom", subtitle =expression("Comparison of the number of species names found in \n analyzed data (local) and in NCBI Taxonomy")) +
  theme_ipsum(
    base_family = 'Arial'
  ) +
  theme(
    panel.spacing = unit(0,'lines'),
    plot.subtitle = element_text(margin = margin(2,0,0,0, "mm")),
    axis.title = element_blank(),
    strip.text = element_text(size=10),
    axis.text.x = element_text(size=6),
    axis.text.y = element_text(size=8),
  )
plt
```

Save plot

```{r}
ggsave('images/taxon-cov-by-kingdom-ncbi.png', plt, dpi = 300)
```

### Calcute coverages

GBIF: by kingdom and overall coverage

```{r}
species.by.kingdom %>%
  mutate(coverage=round(my/gbif*100,1))
species.by.kingdom %>%
  summarise(my=sum(my), gbif=sum(gbif), coverage=round(my/gbif*100,2))
```

NCBI: by kingdom and overall coverage

```{r}
species.ncbi.by.kingdom %>%
  mutate(coverage=round(my/ncbi*100,1))
species.ncbi.by.kingdom %>%
  summarise(my=sum(my), ncbi=sum(ncbi), coverage=round(my/ncbi*100,2))
```

## Completeness Interaction records

Replace empty coordinates and date for `NA`
```{r}
interactions <- interactions %>%
  dplyr::mutate(
    decimalLatitude=replace(decimalLatitude, decimalLatitude=="", NA), 
    decimalLongitude=replace(decimalLongitude, decimalLongitude=="", NA),
    eventDate=replace(eventDate, eventDate=="", NA)
    )
```

Preparing data for counting interactions by taxon rank. First remove interaction without GBIF taxon IDs, then, create flag for records with/without geographic coordinates (`hasloc`) and temporal data (`hasdate`) and both (`haslocdata`).
```{r}
counts.by.rank <- interactions %>%
  filter(stringr::str_starts(sourceTaxonId, "GBIF:") & stringr::str_starts(targetTaxonId, "GBIF:")) %>%
  dplyr::mutate(hasloc=!is.na(decimalLatitude) & !is.na(decimalLongitude), hasdate=!is.na(eventDate), haslocdate=!is.na(decimalLatitude) & !is.na(decimalLongitude) & !is.na(eventDate))
head(counts.by.rank)
```

Counting number of interaction records at species level by kingdom
```{r}
counts.species.rank <- counts.by.rank %>%
  distinct(sourceTaxonKingdomName, sourceTaxonSpeciesName, .keep_all=T) %>%
  group_by(sourceTaxonKingdomName) %>%
  dplyr::summarise(n=n(),nloc=sum(hasloc), ndate=sum(hasdate), nlocdate=sum(haslocdate))
counts.species.rank
```
Counting number of interaction records from higher taxon ranks
```{r}
count.up.species.rank <- counts.by.rank %>%
  distinct(sourceTaxonKingdomName, sourceTaxonPhylumName, sourceTaxonClassName, sourceTaxonOrderName, sourceTaxonFamilyName, sourceTaxonGenusName, .keep_all = T) %>%
  group_by(sourceTaxonKingdomName) %>%
  dplyr::summarise(n=n(),nloc=sum(hasloc), ndate=sum(hasdate), nlocdate=sum(haslocdate))
count.up.species.rank
```

Combine the both data.frame into one data.frame and set the `group` variable (`species` = species level identification of taxa and `other` = higher rank/unranked)

```{r}
counts.by.rank <- counts.species.rank %>%
  mutate(group="species") %>%
  bind_rows(count.up.species.rank %>% mutate(group="other")) %>%
  filter(sourceTaxonKingdomName!="incertae sedis" & sourceTaxonKingdomName!="Viruses")  %>%
  mutate(sourceTaxonKingdomName=as.factor(stringr::str_trim(stringr::str_replace_all(sourceTaxonKingdomName,'"', ""))))
counts.by.rank %>%
  arrange(desc(n))
```

Reorganize data.frame for plot and calculate percentages
```{r}
counts.by.rank <- counts.by.rank %>%
  group_by(sourceTaxonKingdomName) %>%
  dplyr::mutate(tot=sum(n)) %>%
  ungroup() %>%
  dplyr::mutate(np=n/tot*100,nlocp=nloc/tot*100,ndatep=ndate/tot*100, nlocdatep=nlocdate/tot*100) %>%
  dplyr::select(sourceTaxonKingdomName, group, nlocp,ndatep,nlocdatep) %>% 
  complete(sourceTaxonKingdomName,group, fill = list(nlocp=0,ndatep=0,nlocdate=0)) %>% 
  gather(key="measure", value="value", -c(1,2))
counts.by.rank
```

Rename measures for plot
```{r}
counts.by.rank$measure[counts.by.rank$measure == 'nlocp'] <- 'w/ location'
counts.by.rank$measure[counts.by.rank$measure == 'ndatep'] <- 'w/ time'
counts.by.rank$measure[counts.by.rank$measure == 'nlocdatep'] <- 'w/ loc. and time'
```

### Plot 

Set a number of `empty_bar` to add at the end of each group.
```{r}
empty_bar <- 2
nObsType <- nlevels(as.factor(counts.by.rank$group))
to_add <- data.frame( matrix(NA, empty_bar*nlevels(counts.by.rank$sourceTaxonKingdomName)*nObsType, ncol(counts.by.rank)) )
colnames(to_add) <- colnames(counts.by.rank)
to_add$sourceTaxonKingdomName <- rep(levels(counts.by.rank$sourceTaxonKingdomName), each=empty_bar*nObsType )
data <- rbind(counts.by.rank, to_add)
data <- data %>% arrange(sourceTaxonKingdomName, measure)
data$id <- rep( seq(1, nrow(data)/nObsType) , each=nObsType)
```


Get the name and the y position of each label
```{r}
label_data <- data %>% 
  dplyr::group_by(id, measure) %>% 
  dplyr::summarize(tot=sum(value)) %>%
  ungroup()
number_of_bar <- nrow(label_data)
angle <- 90 - 360 * (label_data$id-0.5) /number_of_bar     # I substract 0.5 because the letter must have the angle of the center of the bars. Not extreme right(1) or extreme left (0)
label_data$hjust <- ifelse( angle < -90, 1, 0)
label_data$angle <- ifelse(angle < -90, angle+180, angle)
```

Prepare a data frame for base lines
```{r}
base_data <- data %>% 
  group_by(sourceTaxonKingdomName) %>% 
  dplyr::summarize(start=min(id), end=max(id) - empty_bar) %>% 
  rowwise() %>% 
  dplyr::mutate(title=mean(c(start, end)))
```

Prepare a data frame for grid (scales)
```{r}
grid_data <- base_data
grid_data$end <- grid_data$end[ c( nrow(grid_data), 1:nrow(grid_data)-1)] + 1
grid_data$start <- grid_data$start - 1
grid_data <- grid_data[-1,]
```

Make the plot
```{r}
p <- data %>% 
  ggplot() +      
  # Add the stacked bar
  geom_bar(aes(x=as.factor(id), y=value, fill=group), stat="identity", alpha=0.5) +
  scale_fill_viridis(discrete=TRUE, name="Rank", option = 'D', labels=c("Higher ranks/Unranked", "Species"), alpha = 1) +
  
  # Add a val=100/75/50/25 lines. I do it at the beginning to make sur barplots are OVER it.
  geom_segment(data=grid_data, aes(x = end, y = 0, xend = start, yend = 0), colour = "grey", alpha=1, size=0.3 , inherit.aes = FALSE ) +
  geom_segment(data=grid_data, aes(x = end, y = 20, xend = start, yend = 20), colour = "grey", alpha=1, size=0.3 , inherit.aes = FALSE ) +
  geom_segment(data=grid_data, aes(x = end, y = 40, xend = start, yend = 40), colour = "grey", alpha=1, size=0.3 , inherit.aes = FALSE ) +
  geom_segment(data=grid_data, aes(x = end, y = 60, xend = start, yend = 60), colour = "grey", alpha=1, size=0.3 , inherit.aes = FALSE ) +
  geom_segment(data=grid_data, aes(x = end, y = 80, xend = start, yend = 80), colour = "grey", alpha=1, size=0.3 , inherit.aes = FALSE ) +
  geom_segment(data=grid_data, aes(x = end, y = 100, xend = start, yend = 100), colour = "grey", alpha=1, size=0.3 , inherit.aes = FALSE ) +
  
  # Add text showing the value of each 100/75/50/25 lines
  ggplot2::annotate("text", x = rep(max(data$id),6), y = c(0, 20, 40, 60, 80, 100), label = c("0%", "20%", "40%", "60%", "80%", "100%") , color="grey", size=2 , angle=0, fontface="bold", hjust=1) +
  
  ylim(-80,max(label_data$tot, na.rm=T)+10) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    plot.margin = unit(rep(0.1,4), "cm"),
    plot.title = element_text(
      family = 'Econ Sans Cnd',
      face = 'bold',
      size = 12
    ),
    plot.subtitle = element_text(
      family = 'Econ Sans Cnd',
      size=8
    )
  ) +
  coord_polar() +
  
  # Add labels on top of each bar
  geom_text(data=label_data, aes(x=id, y=tot+10, label=measure, hjust=hjust), color="black", fontface="bold",alpha=0.6, size=1.5, angle= label_data$angle, inherit.aes = FALSE ) +
  
  # Add base line information
  geom_segment(data=base_data, aes(x = start, y = -5, xend = end, yend = -5), colour = "black", alpha=0.8, linewidth=0.6 , inherit.aes = FALSE )  +
  geom_text(data=base_data, aes(x = title, y = c(-20,-27,-27,-27,-20,-27,-27), label=sourceTaxonKingdomName), hjust=c(0.5,0.5,0.6,0.5,0.5,0.5,0.4), colour = viridis(n=nlevels(base_data$sourceTaxonKingdomName)), alpha=0.8, size=2, fontface="bold", inherit.aes = FALSE) +
  labs(
    title="Completeness of biotic interaction records",
    subtitle = 'Percentage of records with location, time and both location and time data'
  )
p
```
Save plot
```{r}
ggsave(p, file="images/completeness.png", dpi=300, bg="white")
```



## Summarizing interactions by kingdom

Normalize kingdom names
```{r}
plants <- c('Plantae', 'Viridiplantae', 'Chloroplastida','Archaeplastida')
animals <- c('Animalia', 'Metazoa', 'Protozoa')
int.kingdoms <- interactions %>%
  mutate(sourceTaxonKingdomName=case_when(
    sourceTaxonKingdomName %in% viruses ~ 'Viruses', 
    sourceTaxonKingdomName %in% plants ~ 'Plantae',
    sourceTaxonKingdomName %in% animals ~ 'Animalia',
    sourceTaxonKingdomName == 'Chromista' ~ 'Fungi',
    sourceTaxonKingdomName == 'Eukaryota' ~ 'incertae sedis',
    .default = sourceTaxonKingdomName),
    targetTaxonKingdomName=case_when(
      targetTaxonKingdomName %in% viruses ~ 'Viruses', 
      targetTaxonKingdomName %in% plants ~ 'Plantae',
      targetTaxonKingdomName %in% animals ~ 'Animalia',
      targetTaxonKingdomName == 'Chromista' ~ 'Fungi',
      targetTaxonKingdomName == 'Eukaryota' ~ 'incertae sedis',
      .default = targetTaxonKingdomName)
    )
dim(int.kingdoms)
```

Count interactions by kingdom

```{r}
int.kingdoms <- int.kingdoms %>%
  dplyr::count(sourceTaxonKingdomName, targetTaxonKingdomName) %>%
  arrange(sourceTaxonKingdomName, targetTaxonKingdomName, n)

int.kingdoms[which(int.kingdoms$sourceTaxonKingdomName > int.kingdoms$targetTaxonKingdomName), c("sourceTaxonKingdomName", "targetTaxonKingdomName")] <-
  rev(int.kingdoms[which(int.kingdoms$sourceTaxonKingdomName > int.kingdoms$targetTaxonKingdomName), c("sourceTaxonKingdomName","targetTaxonKingdomName")])

int.kingdoms <- int.kingdoms %>%
  group_by(sourceTaxonKingdomName, targetTaxonKingdomName) %>%
  dplyr::summarise(n=sum(n))

int.kingdoms %>%
  group_by(sourceTaxonKingdomName, targetTaxonKingdomName) %>%
  arrange(desc(n)) %>%
  mutate(perc=round(n/sum(int.kingdoms$n)*100,1))
```

## Most common interaction by kingdom

```{r}
int.kingdoms <- interactions %>%
  filter(sourceTaxonId != 'no:match' & targetTaxonId != 'no:match') %>%
  filter(sourceTaxonKingdomName!="" & targetTaxonKingdomName!="") %>%
  mutate(sourceTaxonKingdomName=case_when(
    sourceTaxonKingdomName %in% viruses ~ 'Viruses', 
    sourceTaxonKingdomName %in% plants ~ 'Plantae',
    sourceTaxonKingdomName %in% animals ~ 'Animalia',
    sourceTaxonKingdomName == 'Chromista' ~ 'Fungi',
    .default = sourceTaxonKingdomName),
    targetTaxonKingdomName=case_when(
      targetTaxonKingdomName %in% viruses ~ 'Viruses', 
      targetTaxonKingdomName %in% plants ~ 'Plantae',
      targetTaxonKingdomName %in% animals ~ 'Animalia',
      targetTaxonKingdomName == 'Chromista' ~ 'Fungi',
      .default = targetTaxonKingdomName)
    ) %>%
  dplyr::count(sourceTaxonKingdomName, targetTaxonKingdomName, interactionTypeName) %>%
  arrange(sourceTaxonKingdomName, targetTaxonKingdomName, n)

int.kingdoms[which(int.kingdoms$sourceTaxonKingdomName > int.kingdoms$targetTaxonKingdomName), c("sourceTaxonKingdomName", "targetTaxonKingdomName")] <-
  rev(int.kingdoms[which(int.kingdoms$sourceTaxonKingdomName > int.kingdoms$targetTaxonKingdomName), c("sourceTaxonKingdomName","targetTaxonKingdomName")])

int.kingdoms <- int.kingdoms %>%
  group_by(sourceTaxonKingdomName, targetTaxonKingdomName, interactionTypeName) %>%
  dplyr::summarise(n=sum(n)) %>%
  arrange(desc(n))
int.kingdoms
```

### Plants

## Most common plant families
```{r}
interactions %>%
  filter(sourceTaxonKingdomName %in% plants | targetTaxonKingdomName %in% plants) %>%
  dplyr::rename(id=sourceTaxonId,speciesName=sourceTaxonSpeciesName,kingdomId=sourceTaxonKingdomId,kingdomName=sourceTaxonKingdomName, family=sourceTaxonFamilyName) %>%
  select(id,speciesName,kingdomId,kingdomName, family) %>%
  bind_rows(
    interactions %>%
      filter(sourceTaxonKingdomName %in% plants | targetTaxonKingdomName %in% plants) %>%
      dplyr::rename(id=targetTaxonId, speciesName=targetTaxonSpeciesName, kingdomId=targetTaxonKingdomId, kingdomName=targetTaxonKingdomName,family=targetTaxonFamilyName) %>%
      select(id,speciesName,kingdomId,kingdomName, family)
  ) %>%
  filter(kingdomName %in% plants) %>%
  group_by(family) %>%
  dplyr::summarise(n=n(), perc=round(n/(nrow(interactions)*2)*100,1)) %>%
  arrange(desc(n))
  
```

```{r}
int.kingdoms %>%
  filter(sourceTaxonKingdomName=='Plantae' | targetTaxonKingdomName=='Plantae') %>%
  group_by(sourceTaxonKingdomName, targetTaxonKingdomName, interactionTypeName) %>%
  arrange(desc(n)) %>%
  mutate(perc=round(n/sum(int.kingdoms %>% filter(sourceTaxonKingdomName=='Plantae' | targetTaxonKingdomName=='Plantae') %>% pull(n))*100,1)) %>%
  group_by(interactionTypeName) %>%
  dplyr::summarise(n=sum(n), perc=sum(perc)) %>%
  arrange(desc(n))
```
Find the plant - kill/preysOn interactions
```{r}
interactions %>%
  filter(sourceTaxonKingdomName=='Plantae' | targetTaxonKingdomName=='Plantae') %>%
  filter(interactionTypeName=='kills' | interactionTypeName=='preysOn')
```

Number of datasets with `hostOf` interactions:
```{r}

```

### Animals

## Most common plant families
```{r}
interactions %>%
  filter(sourceTaxonKingdomName %in% animals | targetTaxonKingdomName %in% animals) %>%
  dplyr::rename(id=sourceTaxonId,speciesName=sourceTaxonSpeciesName,kingdomId=sourceTaxonKingdomId,kingdomName=sourceTaxonKingdomName, order=sourceTaxonOrderName) %>%
  select(id,speciesName,kingdomId,kingdomName, order) %>%
  bind_rows(
    interactions %>%
      filter(sourceTaxonKingdomName %in% animals | targetTaxonKingdomName %in% animals) %>%
      dplyr::rename(id=targetTaxonId, speciesName=targetTaxonSpeciesName, kingdomId=targetTaxonKingdomId, kingdomName=targetTaxonKingdomName,order=targetTaxonOrderName) %>%
      select(id,speciesName,kingdomId,kingdomName, order)
  ) %>%
  filter(kingdomName %in% animals) %>%
  group_by(order) %>%
  dplyr::summarise(n=n(), perc=round(n/(nrow(.)*2)*100,1)) %>%
  arrange(desc(n))
  
```

```{r}
int.kingdoms %>%
  filter(sourceTaxonKingdomName=='Animalia' | targetTaxonKingdomName=='Animalia') %>%
  group_by(sourceTaxonKingdomName, targetTaxonKingdomName, interactionTypeName) %>%
  arrange(desc(n)) %>%
  mutate(perc=round(n/sum(int.kingdoms %>% filter(sourceTaxonKingdomName=='Animalia' | targetTaxonKingdomName=='Animalia') %>% pull(n))*100,1)) %>%
  group_by(interactionTypeName) %>%
  dplyr::summarise(n=sum(n), perc=sum(perc)) %>%
  arrange(desc(n))
```




