---
title: "Geographic Analysis of biotic interactions data"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages
```{r}
library(dplyr)
#library(tidyverse)
library(maps)
library(ggplot2)
library(rnaturalearth)
library(sf)
library(viridis)
library(scales)
library(RColorBrewer)
library(wdpar)
```

Define some colors for mapping
```{r}
# Colors
BLUE <- "#076fa2"
RED <- "#E3120B"
BLACK <- "#202020"
GREY <- "grey50"
GREEN <- "#90ff01"
GREEN2 <- "#69b3a2"


# Interaction types colors
INT_TYPES_COLORS = c(
  "interactsWith" = "#069668", 
  "hasHost" = "#c02a85", 
  "preysOn" = "#6fed8f",
  "eats" = "#fa79f5",
  "coOccursWith" = "#3aa609",
  "epiphyteOf" = "#9739d9",
  "visitsFlowersOf" =  "#76f014", 
  "muatualistOf" =  "#fe16f4",
  "createdHabitatFor" = "#c5df72",
  "parasiteOf" = "#473c85",
  "kills" = "#f6cf89",
  "ectoparasiteOf" =  "#8f0f1b",
  "symbiontOf" = "#93e0f0"
)
```


# Load and prepate data 

Load interactions with location and time data
```{r}
interactions.data <- read.csv("data/interactions-taxon-loc-time.tsv.gz", header=T, sep = '\t', quote="")
names(interactions.data)
head(interactions.data)
dim(interactions.data)
```

### DEBUG
```{r}
#saveRDS(interactions.data, file = 'data/interactions.data.tmp.rds')
interactions.data <- interactions.data %>%
  sample_n(10000)
head(interactions.data)
```

Remove records with missing taxonomic data:
```{r}
int.copy <- interactions.data
interactions.data <- int.copy %>%
  filter(sourceTaxonId != 'no:match' & targetTaxonId != 'no:match') %>%
  filter(sourceTaxonSpeciesName != "" & targetTaxonSpeciesName != '')
dim(int.copy)
dim(interactions.data)
```

Remove records missing lat/lon
```{r}
interactions.latlon <- interactions.data %>%
  filter(!is.na(decimalLatitude) & !is.na(decimalLongitude) & decimalLatitude!="" & decimalLongitude!="") %>%
  dplyr::count(decimalLatitude,decimalLongitude, interactionTypeName) %>%
  ungroup() %>% 
  arrange(desc(n))
dim(interactions.latlon)
```

```{r}
sum(interactions.latlon$n)
```

Get countries polygons
```{r}
countries <- ne_countries(scale='large', returnclass = 'sf')
```

Clean data: remove records with invalid or zero coordinates, remove records outside countries boundaries
```{r}
interactions.cleaned <- interactions.latlon %>%
  CoordinateCleaner::cc_val(lat="decimalLatitude", lon="decimalLongitude", verbose = F) %>%
  CoordinateCleaner::cc_zero(lat="decimalLatitude", lon="decimalLongitude", verbose=F)
dim(interactions.cleaned)

#interactions.cleaned.sea <- interactions.cleaned %>%
#  CoordinateCleaner::cc_sea(lat="decimalLatitude", lon="decimalLongitude", verbose = F, speedup = T,ref = sf::as_Spatial(countries)) 
```

```{r}
sum(interactions.cleaned$n)
```


Get WDPA (protected areas spatial data)
```{r}
#wdpa <- readRDS('data/wdpa-cleaned.rds')
#wdpa <- wdpa_fetch('global', wait = TRUE, download_dir = 'data/wdpa')
wdpa <- st_read('data/wdpa/wdpa-qgis.shp', stringsAsFactors=F)
```

Clean data WPDA data
```{r}
wdpa <- wdpa_clean(wdpa, erase_overlaps = FALSE, geometry_precision=100)
```

Remove POINTS
```{r}
wdpa <- wdpa %>%
  filter(GEOMETRY_TYPE=='POLYGON')
```

Dissolve small polygons
```{r}
wdpa <- wdpa_dissolve(wdpa, geometry_precision = 100) 
```

Reproject to WGD84
```{r}
wdpa <- st_transform(wdpa, 4326)
wdpa
```

Save RDS
```{r}
#saveRDS(wdpa, file='data/wdpa-clean.rds')
wdpa
```

# DEBUG
```{r}
wdpa <- wdpa %>% sample_n(10000)
```

World boundaries layer
```{r}
world <- map_data('world')
```

# Plots

Plot interaction records 
```{r}
interactions.cleaned  %>%
  ggplot(mapping = aes(x=decimalLongitude, y=decimalLatitude)) +
  geom_point(aes(colour=as.factor(interactionTypeName), size=n))
```


World distribution of interaction observations and WDPA polygons (protected areas)
```{r}
p <- interactions.cleaned  %>%
  ggplot() +
  geom_polygon(data = world, aes(x=long, y = lat, group = group), fill="grey90", alpha=0.8) +
  #geom_hex(bins=200) +
  geom_point(aes(x=decimalLongitude, y=decimalLatitude, colour=interactionTypeName), size=0.5, alpha=1) +
  #geom_sf(data=wdpa, linewidth=0.5, fill="#FEFF00", alpha=0.5, show.legend = F) +
  #scale_size_continuous(
  #  name = "Num. Interactions",
    #trans = 'log10',
  #  range = c(2, 7),
  #  guide = guide_legend(keyheight = unit(2.5, units = "mm"), keywidth=unit(10, units = "mm"), label.position = "bottom", title.position = 'top', nrow=2) 
  #) +
  scale_color_manual(
    values = INT_TYPES_COLORS,
    name="Interaction type", 
    guide = guide_legend(override.aes = list(size = 5), keyheight = unit(2.5, units = "mm"), keywidth=unit(10, units = "mm"), label.position = "bottom", title.position = 'top', nrow=2) 
  )  +
  scale_x_continuous(limits = c(-180,180)) +
  scale_y_continuous(limits = c(-90,90)) +
  theme_void() + theme(
    legend.position = "bottom",
    legend.key.size = unit(3,"cm"),
  )
p
```

Save plot
```{r}
ggsave(filename = 'images/interactions-wdpa-worldmap.pdf', plot=p, dpi=300)
```


```{r}
top <- wdpa %>%
  arrange(desc(REP_AREA))
top
```


Map without interaction of type: interactionWith and coOccursWith
```{r}
p <- interactions.cleaned  %>%
  filter(interactionTypeName!="interactsWith" & interactionTypeName!="coOccursWith") %>%
  ggplot() +
  geom_sf(data=wdpa, linewidth=NA, fill="#c5df72", color=NA, alpha=0.25, show.legend = F) +
  geom_polygon(data = world, aes(x=long, y = lat, group = group), fill="grey90", colour=NA, linewidth=NA, alpha=0.3) +
  #geom_hex(bins=200) +
  geom_point(aes(x=decimalLongitude, y=decimalLatitude, colour=interactionTypeName), size=0.5, alpha=1) +
  #scale_size_continuous(
  #  name = "Num. Interactions",
    #trans = 'log10',
    #range = c(2, 7),
  #  guide = guide_legend(keyheight = unit(2.5, units = "mm"), keywidth=unit(10, units = "mm"), label.position = "bottom", title.position = 'top', nrow=2) 
  #) +
  scale_color_manual(
    values = INT_TYPES_COLORS,
    name="Interaction type", 
    guide = guide_legend(override.aes = list(size = 5), keyheight = unit(2.5, units = "mm"), keywidth=unit(10, units = "mm"), label.position = "bottom", title.position = 'top', nrow=2) 
  )  +
  scale_x_continuous(limits = c(-180,180)) +
  scale_y_continuous(limits = c(-90,90)) +
  theme_void() + theme(
    legend.position = "bottom",
    legend.key.size = unit(3,"cm"),
  )
p
```

Save plot
```{r}
ggsave(filename = 'images/interactions-worldmap-wo-generic-inttypes.png', plot=p, dpi=300)
```

## Join Protected Areas and Interactions records

Create a interaction points layer in WGS84
```{r}
int.points <- st_as_sf(interactions.cleaned, coords = c('decimalLongitude','decimalLatitude'), crs = 4326) #"+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")
```

Find intersection between WDPA polygons and interaction points
```{r}
int.wdpa <- st_intersects(int.points, wdpa)
int.points.intersect <- lengths(int.wdpa) > 0
wdpa.intersect <- lengths(t(int.wdpa)) > 0
```

Total area of WDPA
```{r}
sum(wdpa$REP_ARE)
```

Display some summary results
```{r}
print(paste0("% of points inside protected areas: ", round(nrow(int.points[int.points.intersect,])/nrow(int.points)*100,2), '%'))
wdpa.total_area <- sum(wdpa$REP_ARE)
wdpa.total_area
int.total_area <- sum(wdpa[wdpa.intersect,]$REP_ARE)
int.total_area
print(paste("% Total area of protected areas with interaction records:", int.total_area/wdpa.total_area*100))
print(paste("Total area (milions of km2) of protected areas with interaction records:", round(int.total_area,2)))
```


```{r}
st_write(wdpa %>% select(-WDPAID), 'data/wdpa-clean.shp', append=F)
```

World distribution of interaction observations and WDPA polygons (protected areas)
```{r}
p <- interactions.cleaned[int.points.intersect,]  %>%
  ggplot() +
  geom_polygon(data = world, aes(x=long, y = lat, group = group), fill="grey90", alpha=0.8) +
  #geom_hex(bins=200) +
  geom_sf(data=wdpa, linewidth=NA, fill="#c5df72", alpha=0.5, show.legend = F) +
  geom_point(aes(x=decimalLongitude, y=decimalLatitude, colour=interactionTypeName), size=0.2, alpha=1) +
  #scale_size_continuous(
  #  name = "Num. Interactions",
    #trans = 'log10',
  #  range = c(2, 7),
  #  guide = guide_legend(keyheight = unit(2.5, units = "mm"), keywidth=unit(10, units = "mm"), label.position = "bottom", title.position = 'top', nrow=2) 
  #) +
  scale_color_manual(
    values = INT_TYPES_COLORS,
    name="Interaction type", 
    guide = guide_legend(override.aes = list(size = 5), keyheight = unit(2.5, units = "mm"), keywidth=unit(10, units = "mm"), label.position = "bottom", title.position = 'top', nrow=2) 
  )  +
  scale_x_continuous(limits = c(-180,180)) +
  scale_y_continuous(limits = c(-90,90)) +
  theme_void() + theme(
    legend.position = "bottom",
    legend.key.size = unit(3,"cm"),
  )
p
```

Save plot
```{r}
ggsave(filename = 'images/interactions-wdpa-intersection-worldmap.png', plot=p, dpi=300)
```

## Intersection Bioregions and interaction observations

```{r}
library(geodat)
data(zoorealms)
zoorealms <- st_wrap_dateline(zoorealms, options = c("WRAPDATELINE=YES","DATELINEOFFSET=180"))
```

Re-project realm data
```{r}
zoorealms <- sf::st_transform(zoorealms, 4326)
```

Plot map of realm data
```{r}
p <- ggplot() + geom_sf(data=zoorealms, aes(fill=Realm)) + 
  scale_fill_manual(values=ggsci::pal_d3("category20")(11)) + 
  scale_x_continuous(limits = c(-180,180)) +
  scale_y_continuous(limits = c(-90,90)) +
  coord_sf() + 
  theme_bw() + theme(panel.border = element_blank(),
                     legend.title = element_text(size=12, face="bold"))
p
```

Intersection interactions and zoorealms
```{r}
zoorealms <- zoorealms %>% st_make_valid()
zoorealms
int.zooreamls <- st_intersects(int.points, zoorealms, sparse = F)
int.zooreamls
```

Intersection by zoorealms
```{r}
int.by.realm <- apply(int.zooreamls, 2, sum)
names(int.by.realm) <- zoorealms$Realm
total = sum(int.by.realm)
```

```{r}
print("% of records by biogegraphy realm: ")
print(names(int.by.realm))
print(paste(int.by.realm, paste0(round(int.by.realm/total*100, 2), "%"), sep ="/"))
int.by.realm
round(int.by.realm/total*100, 2)
```

Plot interactions over biorealms:
```{r}
```

# Temporal Analysis


Convert `eventDate` to year
```{r}
int.dates <- interactions.data %>%
  filter(eventDate != "")

int.by.year <- int.dates %>%
  mutate(year=as.integer(format(as.Date(eventDate), "%Y")))
max(int.by.year$year, na.rm=T)
```

Filter invalid dates
```{r}
int.by.year <- int.by.year %>%
  filter(year>=1800 & year<=2023)
dim(int.by.year)
```

Plot interactions by year
```{r}
dplyr::count(year, interactionTypeName) %>%
  ungroup() %>%
  arrange(year)
dim(int.by.year)  
int.by.year

plt <- int.by.year %>%
  mutate(year=as.integer(year)) %>%
  ggplot(aes(x=year, y=n, color=interactionTypeName)) +
  geom_point() +
  theme_ipsum(
    base_family = 'Arial',
  ) + 
  theme(
    legend.position = 'bottom'
  )
plt

```
